Reproducible Research Course Project 1
======================================================

```{r}
library(dplyr)
```

Part 1
========================================================

load the data
``` {r load_data}
activity_data <- read.csv("activity.csv")
head(activity_data)
str(activity_data)
```

total number of steps taken per day
```{r total_steps}
total_steps <- with(activity_data, tapply(steps, date, sum, na.rm = T))
head(total_steps)
```

histogram of total steps
```{r hist}
hist(total_steps, col = "blue")
```

mean steps taken each day
```{r mean_steps}
mean_steps <- with(activity_data, tapply(steps, date, mean, na.rm = T))
head(mean_steps)
```

median steps taken each day
```{r median_steps}
median_steps <- with(activity_data, tapply(steps, date, median, na.rm = T))
head(median_steps)
```

Part 2
==========================================

Time series plot of intervals and mean steps
```{r time_series}
int_ave_steps <- with(activity_data, tapply(steps, as.factor(interval), mean, na.rm = T))
plot(names(int_ave_steps), int_ave_steps, type = "l", 
     col = "blue", xlab = "Interval", ylab = "Average Number of Steps", main = "Average Number of Steps Taken Each Interval of a Day")
```

which interval has the max average number of steps
```{r max_step}
which.max(int_ave_steps)
```

Part 3
============================================

number of missing values
```{r missing}
total_na <- sum(is.na(activity_data$steps))
total_na
```

fill in the missing values with mean of intervals
```{r fill}
mean_step_interval <- activity_data %>% group_by(interval) %>% summarise(mean_step = mean(steps, na.rm = T))
activity_data_new <- merge(activity_data, mean_step_interval, by = "interval", all.x = T)
activity_data_new$steps[is.na(activity_data_new$steps)] <- activity_data_new[is.na(activity_data_new$steps), ]$mean_step
```

new data set with missing data filled in
``` {r data:_process}
activity_data_new <- arrange(activity_data_new, date, interval)
activity_data_new <- activity_data_new %>% select(-mean_step)
head(activity_data_new)
```

total number of steps taken per day - new data
```{r total_steps_new}
total_steps_new <- with(activity_data_new, tapply(steps, date, sum, na.rm = T))
head(total_steps_new)
```

histogram of total steps - new data
```{r hist_new}
hist(total_steps_new, col = "blue")
```

mean steps - new data
```{r mean_steps_new}
mean_steps_new <- with(activity_data_new, tapply(steps, date, mean, na.rm = T))
head(mean_steps_new)
```

median steps - new data
```{r median_steps_new}
median_steps_new <- with(activity_data_new, tapply(steps, date, median, na.rm = T))
head(median_steps_new)
```

Part 4
===================================================

weekday/weekend
```{r weekday}
activity_data_new <- activity_data_new %>% mutate(day = weekdays(as.Date(date)))
activity_data_new <- activity_data_new %>% mutate(day = if_else(day %in% c("Sunday", "Saturday"), "weekend", "weekday"))
head(activity_data_new)
```

Time series plot of intervals and mean steps for weekday/weekend
```{r panel_plot}
mean_step_interval_new <- activity_data_new %>% group_by(interval) %>% summarise(mean_step = mean(steps, na.rm = T))
activity_data_new <- merge(activity_data_new, mean_step_interval_new, by = "interval", all.x = T)
par(mfrow = c(2,1)) 
with(subset(activity_data_new, day = "weekday"), plot(interval, mean_step, type = "l", main = "Weekdays", col = "purple"))
with(subset(activity_data_new, day = "weekend"), plot(interval, mean_step, type = "l", main = "weekend", col = "purple"))
``` 
